---
title: ""
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 5
---

<br>

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #337ab7;
}

.navbar-default .navbar-collapse, .navbar-default .navbar-form {
background-color: #337ab7;
}

#Drop-Down Menu Color
.navbar-default .navbar-nav>.open>a, .navbar-default .navbar-nav>.open>a:focus, .navbar-default .navbar-nav>.open>a:hover {
background-color: #337ab7;
}

.navbar-default {
background-color: #337ab7;
}

.navbar-default .navbar-nav>li>a {
color: white;
font-weight: bold;
}

.navbar-default .navbar-brand {
color: white;
font-weight: bold;
}

</style>

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

##Exploring NYC Crime Data Using EDA - Cleaning Data

```{r setup, include = FALSE}
knitr::opts_chunk$set(
 echo = TRUE,
 out.width = "100%"
)
```

<br>

###Importing Data

Below, we include the code we used to import and clean the data for the "Exploring NYC Crime Data Using EDA" project:

```{r load_packages}
library(RSocrata)
library(tidyverse)
library(lubridate)
```


```{r import_data, eval = FALSE}
nyc_crime <- read_csv("./data/NYPD_Complaint_Data_Historic.csv")

#Code for importing directly from web:
#nyc_crime = read.socrata("https://data.cityofnewyork.us/resource/9s4h-37hy.json", 
#                          app_token = NULL, 
#                          email = NULL, 
#                          password = NULL,
#                          stringsAsFactors = FALSE)

#saveRDS(nyc_crime, file = "./data/nyc_crime.rds")
```

The raw, acquired dataset has 6,036,805 observations and 35 variables. Broadly speaking, the variables contain information on the exact date, time and location of crime, description of crime, demographic information of the victim and suspect, and police department infromation. For more information on the variables, click [here](https://data.cityofnewyork.us/Public-Safety/NYPD-Complaint-Data-Historic/qgea-i56i).

<br>

###Penal Codes 

We were interested in sex-related, weapon-related and drug-related felonies occurring in NYC from 2014 - 2017. Below is a list of the penal codes associated with each type of felony.

<br>

####Sex–Related Felonies <br>
`178` Facilitating A Sex Offense With A Controlled Substance <br>
`694` Incest <br>
`697` Use Of A Child In Sexual Performance <br>
`176` Sex Crimes <br>
`180` Course Of Sexual Conduct Against Child <br>
`153` Rape 3 <br>
`157` Rape 1 <br>
`177` Sexual Abuse <br>
`168` Sodomy 1 <br>
`159` Rape 1 Attempt <br>
`166` Sodomy 2 <br>
`164` Sodomy 3 <br>
`179` Aggrevated Sexual Abuse <br>
`155` Rape 2 <br>
`586` Sextrafficking <br>
`696` Promoting Sexual Performance – Child 

<br>

####Drug-Related Felonies <br>

`500` Controlled Substance, Possession <br>
`501` Controlled Substance, Possession <br>
`502` Controlled Substance, Possession <br>
`503` Controlled Substance, Intent To <br>
`505` Controlled Substance, Possession <br>
`507` Controlled Substance, Possession <br>
`510` Controlled Substance, Intent T <br>
`512` Controlled Substance, Sale 1 <br>
`514` Controlled Substance, Sale 2 <br>
`515` Controlled Substance, Sale 3 <br>
`519` Sale School Grounds 4 <br>
`520` Controlled Substance, Sale 4 <br>
`521` Controlled Substance, Sale 5 <br>
`523` Sale School Grounds <br>
`524` Controlled Substance, Possession<br>
`529` Sales Of Prescription <br>
`530` Drug, Injection Of <br>
`531` Drug Paraphernalia, Possessesion <br>
`532` Controlled Substance,Possession <br>
`568` Marijuana, Possession 1, 2 & 3 <br>
`570` Marijuana, Sale 1, 2 & 3 

<br>

####Weapons-Related Felonies <br> 

`781` Criminal Disposal Firearm 1 <br>
`792` Weapons Possession 1 & 2 <br>
`793` Weapons Possession 3 <br>
`796` Weapons,Prohibited Use 

<br>

###Subsetting Data <br>

Using that information, we can subset and filter the data to provide us with our final dataset.
```{r subset_data, eval = FALSE}
#Cleaning and Filtering Data
nyc_felonies = nyc_crime %>% 
  janitor::clean_names() %>% 
  mutate(year = year(cmplnt_fr_dt)) %>%
  mutate_if(is.character, tolower) %>% 
  filter(year %in% 2014:2017) %>% 
  filter(law_cat_cd == "felony") %>% 
  select(- station_name, - transit_district, - hadevelopt, - patrol_boro, - housing_psa, 
         - juris_desc)

#saveRDS(nyc_felonies, file = "./data/nyc_felonies.rds")

#Selecting Crimes of Interest
sex_drug_weapons = nyc_felonies %>% 
  
  filter(pd_cd %in% c(178, 694, 697, 176, 180, 153, 157, 177, 168, 159, 166, 164, 179, 155, 
                      586, 696, # Sex-related felonies
                      
                      ## Drug-related felonies
                      500, 501, 502, 503, 505, 507, 510, 512, 514, 515, 519, 520, 521, 523, 
                      524, 529, 530, 531, 532, 568, 570,
                      
                      ### Weapons-related felonies
                      781, 792, 793, 796)) %>% 
  
  #Select Variables of Interest
  select(cmplnt_num, boro_nm, cmplnt_fr_dt, cmplnt_to_dt, cmplnt_fr_tm, ky_cd, ofns_desc, 
         pd_cd, pd_desc, vic_race, vic_sex, vic_age_group, year, prem_typ_desc) %>% 
  
  #Create Classification of Felonies
  mutate(boro_nm = if_else(boro_nm == "staten island", "staten_island", boro_nm),
         
     crime_group = if_else(pd_cd %in% c(178, 694, 697, 176, 180, 153, 157, 177, 168, 159, 166, 
                                        164, 179, 155, 586, 696), "Sex-Related", 
                               
                   if_else(pd_cd %in% c(500, 501, 502, 503, 505, 507, 510, 512, 514, 515, 519, 
                                        520, 521, 523, 524, 529, 530, 531, 532, 568, 570), 
                                        "Drug-Related", 
                               
                   if_else(pd_cd %in% c(781, 792, 793, 796), "Weapons-Related", pd_cd))))

#saveRDS(sex_drug_weapons, file = "./data/sex_drug_weapons.rds")
```

<br>

The resulting dataset has **46,692 observations** and **14 variables**.
```{r final_dataset, eval = FALSE}
felonies = readRDS(file = "./data/sex_drug_weapons.rds")

summary(felonies)
```

