---
title: ""
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 5
---

<br>

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #337ab7;
}

.navbar-default .navbar-collapse, .navbar-default .navbar-form {
background-color: #337ab7;
}

#Drop-Down Menu Color
.navbar-default .navbar-nav>.open>a, .navbar-default .navbar-nav>.open>a:focus, .navbar-default .navbar-nav>.open>a:hover {
background-color: #337ab7;
}

.navbar-default {
background-color: #337ab7;
}

.navbar-default .navbar-nav>li>a {
color: white;
font-weight: bold;
}

.navbar-default .navbar-brand {
color: white;
font-weight: bold;
}

</style>

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

##SQL Queries With NHANES Data

<br>

###Skills 

* SQL

```{r setup, include = FALSE}
knitr::opts_chunk$set(
 echo = TRUE,
 out.width = "100%"
)

library(tidyverse)
library(sqldf)
```

<br>

###Task

Complete the following questions using SQL.

<br>

###Data

NHANES data containing demographic information for respondents can be found [here](data/NHANES_Demographics.csv). NHANES data containing additional information on respondents, such as triglyceride and cholesterol levels, can be found [here](data/NHANES_Triglycerides). Relevant information on the variables is included below:

<p> `Gender` - Gender of respondents <br> <p>
* 1 - Male
* 2 - Female <br>

<p> `Age_in_years_at_screening` = Age at screening in years (continuous up to 79) <br> <p>
* 0-79 - Any age between 0 and 79
* 80 - 80 or older <br>

<p> `Race_Hispanic_origin_w_NH_Asian` = Race/Hispanic origin <br> <p>
* 1 - Mexican American
* 2 - Other Hispanic
* 3 - Non-Hispanic White
* 4 - Non-Hispanic Black
* 6 - Non-Hispanic Asian
* 7 - Other race, including multi-racial
* NA - Missing <br>

<p> `Marital_status` = Marital status  <br> <p>
* 1 - Married
* 2 - Widowed
* 3 - Divorced
* 4 - Separated
* 5 - Never married
* 6 - Living with partner
* 77 - Refused
* 99 - Don't know
* NA - Missing <br>

<p> `Annual_household_income` = Annual household income <br> <p>
* 1 - $0 to $4,999
* 2 - $5,000 to $9,999
* 3 - $10,000 to $14,999
* 4 - $15,000 to $19,999
* 5 - $20,000 to $24,999
* 6 - $25,000 to $34,999
* 7 - $35,000 to $44,999
* 8 - $45,000 to $54,999
* 9 - $55,000 to $64,999
* 10 - $65,000 to $74,999
* 11 - $20,000 and over
* 12 - Under $20,000
* 14 - $75,000 to $99,999
* 15 - $100,000 and over
* 77 - Refused
* 99 - Don't know
* NA - Missing <br>

<p> `Pregnancy_status_at_exam` = Pregnancy status at exam  <br> <p>
* 1 - Yes
* 2 - No
* 3 - Cannot ascertain
* NA - Missing <br>

<p> `Education_level_Adults` = Eucation level <br> <p>
* 1 - Less than 9th grade
* 2 - 9-11th grade (Includes 12th grade with no diploma)
* 3 - High school graduate/GED or equivalent
* 4 - Some college or AA degree
* 5 - College graduate or above
* 7 - Refused
* 9 - Don't know
* NA - Missing <br>

I first import and merge the datasets before running any queries.
```{r import_data}
demo <- read.csv('data/NHANES_Demographics.csv')
tri <- read.csv('data/NHANES_Triglycerides.csv')

merged <- sqldf(
"SELECT *
FROM demo AS d
LEFT JOIN tri AS t
ON d.Respondent_sequence_number = t.Respondent_sequence_number"
)

knitr::kable(head(merged[1:5]))
knitr::kable(head(merged[6:10]))
knitr::kable(head(merged[11:14]))
```

<br>

I rename the variables in the merged dataset and remove the duplicate ID column.
```{r rename_var}
full.data <- sqldf(
"SELECT Respondent_sequence_number      AS id, 
        Gender                          AS gender,
        Age_in_years_at_screening       AS age, 
        Race_Hispanic_origin_w_NH_Asian AS race,
        Education_level_Adults          AS educ, 
        Marital_status                  AS marital_status,
        Pregnancy_status_at_exam        AS pregnancy_status,      
        Annual_household_income         AS household_income,
        Annual_family_income            AS family_income,
        Triglyceride_mg_dL              AS tri_mg_dl,
        Triglyceride_mmol_L             AS tri_mmol_l,
        LDL_cholesterol_mg_dL           AS chol_ldl,
        LDL_cholesterol_mmol_L          AS chol_mmol_l
FROM merged"
)
```


###Queries

1) Write a query that gives the frequency of each race, as well as the average age of participants by race.
```{r race_freq}
race_freq <- sqldf(
"SELECT race, COUNT(race) AS freq, ROUND(AVG(age),1) AS mean_age
  FROM full.data 
  GROUP BY race"
)

knitr::kable(head(race_freq))
```

<br>

2) Show the distribution of race by gender and display all unique race-gender combinations from highest to lowest frequency.  
```{r race_freq2}
race_freq2 <- sqldf(
"SELECT Gender, Race_Hispanic_origin_w_NH_Asian, COUNT(Race_Hispanic_origin_w_NH_Asian) AS freq
    FROM demo 
GROUP BY Race_Hispanic_origin_w_NH_Asian, Gender 
ORDER BY freq DESC"
)

knitr::kable(head(race_freq2))
```

<br>

3) How many women were pregnant at the time of screening?
```{r preg_freq}
preg_freq <- sqldf(
"SELECT COUNT(Pregnancy_status_at_exam) AS preg_at_screen
    FROM demo GROUP BY Pregnancy_status_at_exam
HAVING Pregnancy_status_at_exam = 1"
)

preg_freq
```

table3 <- sqldf("SELECT COUNT(Pregnancy_status_at_exam) AS preg_at_screen
                FROM demo GROUP BY Pregnancy_status_at_exam
                HAVING Pregnancy_status_at_exam = 1")
table3

    
4. How many men refused to provide annual household income?
table4 <- sqldf("SELECT COUNT(Annual_household_income) AS num_refused
                FROM demo 
                WHERE Gender = 1 and Annual_household_income = 77")
table4


5. What is the mean LDL level (mg/dL) for men and women?  Use column alias mean_ldl and round results to 
1 decimal place.  
table5 <- sqldf("SELECT Gender, ROUND(AVG(LDL_cholesterol_mg_dL), 1) AS mean_ldl
                FROM demo AS d
                LEFT JOIN tri AS t
                ON d.Respondent_sequence_number = t.Respondent_sequence_number
                GROUP BY Gender")
table5

    
6. Display the minimum and maximum triglyceride levels (mmol/L) for each race.  Use column aliases min_tri and max_tri.

table6 <- sqldf("SELECT Race_Hispanic_origin_w_NH_Asian, MIN(LDL_cholesterol_mg_dL), MAX(LDL_cholesterol_mg_dL)
                FROM demo AS d
                LEFT JOIN tri AS t
                ON d.Respondent_sequence_number = t.Respondent_sequence_number
                GROUP BY Race_Hispanic_origin_w_NH_Asian")
table6

  
