---
title: ""
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 5
---

<br>

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #337ab7;
}

.navbar-default .navbar-collapse, .navbar-default .navbar-form {
background-color: #337ab7;
}

#Drop-Down Menu Color
.navbar-default .navbar-nav>.open>a, .navbar-default .navbar-nav>.open>a:focus, .navbar-default .navbar-nav>.open>a:hover {
background-color: #337ab7;
}

.navbar-default {
background-color: #337ab7;
}

.navbar-default .navbar-nav>li>a {
color: white;
font-weight: bold;
}

.navbar-default .navbar-brand {
color: white;
font-weight: bold;
}

</style>

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

##SQL Queries With Cancer Data

<br>

###Skills 

* SQL

```{r setup, include = FALSE}
knitr::opts_chunk$set(
 echo = TRUE,
 out.width = "100%"
)

library(tidyverse)
library(sqldf)
library(lubridate)
```

<br>

###Task

Complete the following questions using SQL.

<br>

###Data

The following data provide information regarding breast and/or colon cancer patients from a clinic, who have been given drugs (A, B, C, and/or D). Data containing patient diagnoses can be found [here](data/diagnosis.csv). Data containing patient drug treatments can be found [here](data/treatment.csv). Relevant information on the variables is included below:

<p> `patient_id` - Patient ID <br> <p>
`diagnosis_date` - Date of Cancer Diagnosis <br> <p>
`diagnosis_code` - Diagnosis Code (ICD-9-CM) <br> <p>
`diagnosis` - Breast Cancer or Colon Cancer <br> <p>
`treatment_date` - Date Patient Initiated Drug Regimen  <br> <p>
`drug_code` - Drug Regimen (A, B, C, or D) 

<br>

I first import and merge the datasets before running any queries.
```{r import_data}
diag <- read.csv('data/diagnosis.csv')
txt <- read.csv('data/treatment.csv')

merged <- sqldf(
"SELECT d.patient_id, diagnosis_date, diagnosis_code, diagnosis, treatment_date, drug_code
    FROM diag AS d
    INNER JOIN txt AS t
    ON d.patient_id = t.patient_id"
)

knitr::kable(head(merged[1:3]))
knitr::kable(head(merged[4:6]))
```

<br>

###Queries

####Q1
Write a query that finds the number of unique patients.

**There are 46 unique patients.**
```{r unique_patients}
unique_patients <- sqldf(
"SELECT COUNT(DISTINCT patient_id) AS unique_patients
    FROM merged" 
)

unique_patients
#46 unique patients
```

<br>

####Q2

How many patients have breast cancer only, colon cancer only, and both breast and colon cancer?

**We have 30 patients breast cancer only patients, 11 colon cancer only patients, and 5 patients with both cancers.**
```{r cancer_num}
breast <- sqldf(
"SELECT COUNT(DISTINCT patient_id)
    FROM merged
WHERE diagnosis = 'Breast Cancer'"
)
breast
#35 patients with breast cancer
```

```{r cancer_num2}
colon <- sqldf(
"SELECT COUNT(DISTINCT patient_id)
    FROM merged
WHERE diagnosis = 'Colon Cancer'"
) 
colon
#16 patients with colon cancer.
```

```{r comment}
#Since there are only 46 unique patients, and 35+16 = 51, this suggests we have 5 individuals who have both cancers.
#This means we have 35 - 5 = 30 breast cancer only patients and 16 - 5 = 11 colon cancer only patients
```

```{r both_cancers}
both_cancers <- sqldf(
"SELECT DISTINCT patient_id, diagnosis
    FROM merged
ORDER BY patient_id")
both_cancers
#Code above confirms that we have 5 patients with both cancers: IDs 3757, 6877, 6889, 6922, and 7230
#Code also confirms numbers for breast cancer only patients and colon cancer only patients
```

<br>

####Q3

What's the average duration between diagnosis date and the earliest treatment date among breast cancer patients? Among colon cancer patients?

**On average, colon cancer patients tend to start treatment 22.9375 days after being diagnosed while breast cancer patients tend to start treatment 4.857143 days after being diagnosed.**
```{r duration1}
merged$diagnosis_date <- as.Date(merged$diagnosis_date, format("%m/%d/%y"))
merged$treatment_date <- as.Date(merged$treatment_date, format("%m/%d/%y"))

breast_dur <- sqldf(
"SELECT AVG(min_date_diff)
    FROM
    (SELECT patient_id, MIN(treatment_date - diagnosis_date) AS min_date_diff
        FROM merged
    WHERE diagnosis = 'Breast Cancer'
    GROUP BY patient_id)"
)
breast_dur
#4.857143 days
```

```{r duration2}
colon_dur <- sqldf(
"SELECT AVG(min_date_diff)
    FROM
    (SELECT patient_id, MIN(treatment_date - diagnosis_date) AS min_date_diff
        FROM merged
    WHERE diagnosis = 'Colon Cancer'
    GROUP BY patient_id)"
)
colon_dur
#22.9375 days
```

```{r comment2}
#Patients with both cancers are included in the calculations for both breast_dur and colon_dur
```

<br>

####Q4

Do breast cancer patients vary in terms of average duration between diagnosis date and the earliest treatment date?
Stratify by drug.

Average duration between diagnosis and treatment for breast cancer patients taking: <p>
* Drug A - 4.782609 days
* Drug B - 5.4 days
* Drug C - 42.13333 days
* Drug D - NA (seems like no breast cancer patients took drug D)
```{r stratify_duration}
breast_a <- sqldf(
"SELECT AVG(min_date_diff)
    FROM
    (SELECT patient_id, MIN(treatment_date - diagnosis_date) AS min_date_diff
        FROM merged
    WHERE diagnosis = 'Breast Cancer' AND drug_code = 'A'
    GROUP BY patient_id)"
)
breast_a


breast_b <- sqldf(
"SELECT AVG(min_date_diff)
    FROM
    (SELECT patient_id, MIN(treatment_date - diagnosis_date) AS min_date_diff
        FROM merged
    WHERE diagnosis = 'Breast Cancer' AND drug_code = 'B'
    GROUP BY patient_id)"
)
breast_b


breast_c <- sqldf(
"SELECT AVG(min_date_diff)
    FROM
    (SELECT patient_id, MIN(treatment_date - diagnosis_date) AS min_date_diff
        FROM merged
    WHERE diagnosis = 'Breast Cancer' AND drug_code = 'C'
    GROUP BY patient_id)"
)
breast_c


breast_d <- sqldf(
"SELECT AVG(min_date_diff)
    FROM
    (SELECT patient_id, MIN(treatment_date - diagnosis_date) AS min_date_diff
        FROM merged
    WHERE diagnosis = 'Breast Cancer' AND drug_code = 'D'
    GROUP BY patient_id)"
)
breast_d
#No breast cancer patients using drug D, so no average duration 
```

```{r comment3}
#Code looks at any drug taken at any point during their treatment dates
```

<br>

####Q5

What tends to be the most popular drug used by breast cancer patients? What about colon cancer patients?

*A and B tend to be the most popular drugs used by breast cancer patients. B and C tend to be the most popular drugs used by colon cancer patients.*
```{r popular_drugs1}

breast_drug <- sqldf(
"SELECT drug_code, COUNT(drug_code)
    FROM
    (SELECT DISTINCT patient_id, drug_code
        FROM merged
    WHERE diagnosis = 'Breast Cancer'
    ORDER BY patient_id)
GROUP BY drug_code"
)
knitr::kable(head(breast_drug))
#A and B tend to be the most popular drugs used by breast cancer patients.
#Observations total more than 35 because some breast cancer patients use multiple drugs.
#Code looks at any drug taken at any point during their treatment dates
```

```{r popular_drugs2}
colon_drug <- sqldf(
"SELECT drug_code, COUNT(drug_code)
    FROM
    (SELECT DISTINCT patient_id, drug_code
        FROM merged
    WHERE diagnosis = 'Colon Cancer'
    ORDER BY patient_id)
GROUP BY drug_code"
)
knitr::kable(head(colon_drug))
#B and C tend to be the most popular drugs used by colon cancer patients.
#Observations total more than 16 because some colon cancer patients use multiple drugs.
#Code looks at any drug taken at any point during their treatment date
```


